---
title: "Figure Generation"
author: "J Fadum"
date: "2023-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}

#PACKAGES
library(tidyverse)
library(signal)
library(xts)
library(dygraphs) 
library(RcppRoll)
library(imputeTS)
library(ggthemes)
library(broom)
library(tsibble)
library(GGally)
library(tidyr)
library(lubridate)
library(readr)
library(plotly) 
library(ggthemes)

#Figure colors...

#Raices #8C510A
#Helado #BF812D
#Varsovia  #01665E
#Yure    #542788 
#Balas  #35978F
#Cacao #8073AC
```

## Figure 2
Note: First call the q_preds dataframe created in DischargeCalcs.Rmd.
We will also load the discharge data for the two tributaries which used Manning's equation (where sensor installation was not possible)

```{r warning=FALSE, error=FALSE, message=FALSE}
cacao_q<- read_csv("q_cacao_2019.csv")
cacao_q$date<- mdy(cacao_q$date)

yure_q<- read_csv("q_yure_2019.csv")
yure_q$date<- mdy(yure_q$date)

q_daily <- q_preds %>%
  inner_join(raw_cm %>% 
               select(site,datetime,temp)) %>%
  group_by(site,date) %>%
  summarize(q = median(med_q),
            count = n(),
            temp = median(temp)) %>%
  dplyr::filter(count > 4*20)

q_daily<- q_daily %>% 
  select(date, q, site)

cacao_q<- cacao_q %>% 
  mutate(site= "cacao") #add site so can combine data sets

yure_q<- yure_q %>% 
  mutate(site= "yure") #add site so can combine data sets

q_daily<- rbind(cacao_q, yure_q, q_daily)  

low_flow<- q_daily %>%     #low flow subset for figure 2
  dplyr::filter(site == "balas"|
                site == "cacao"|
                site == "raices") %>% 
  mutate(cat= "low flow")

high_flow<- q_daily %>%    #higher flow subset for figure 2
  dplyr::filter(site == "helado"|
                site == "varsovia"|
                site == "yure") %>% 
  mutate(cat= "high flow")

low_trib_discharge<- ggplot(low_flow, aes(x = date, y = q, color=site)) + 
  theme_few()+
  geom_line(aes(linetype= site)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Tributary Discharge")+
  ylab("Q (cubic m/sec)")+
  scale_x_date(limits = as.Date(c("2019-04-01", "2020-04-01")))+
   scale_color_manual(values=c("#35978F", "#8073AC", "#8C510A"))

low_trib_discharge

  
high_trib_discharge<- ggplot(high_flow, aes(x = date, y = q, color=site)) + 
  theme_few()+
  geom_line(aes(linetype= site)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Tributary Discharge")+
  ylab("Q (cubic m/sec)")+
  scale_x_date(limits = as.Date(c("2019-04-01", "2020-04-01")))+
   scale_color_manual(values=c("#BF812D", "#01665E", "#542788"))

high_trib_discharge

#LUBRIDATE ISNT WORKING HERE 
rain <- read_csv("precip.csv")
rain$date <- mdy(rain$date) #correct date format with lubridate
str(rain)
precip<- rain%>%  
  mutate(month=month(date)) %>% 
  drop_na() %>% 
  group_by(month)%>%
  summarise(total=sum(rain))

precip$month<- as.character(precip$month)
month_ordered<- ordered(precip$month, levels = c("4", "5", "6", "7", "8", "9", "10", "11", "12", "1", "2", "3")) #order boxplots

precip_plot<- ggplot(precip, aes(y=total, x=month_ordered)) +
  geom_bar(position="stack", stat="identity")+
    ggtitle("Weekly cumulative rain in 2019 (mm)") +
  theme_few()+
  ylab("Precipitation (mm)")+
  xlab("Week")
  
precip_plot
  
```

Now for the rest of the figures we need the nutrient data.

## Figure 3
```{r warning=FALSE, error=FALSE,  message=FALSE}
nutes <- read_csv('13OCT21_Rivers_JFedits.csv') %>% 
  mutate(datetime = mdy_hms(paste(date,'12:00:00')),
         date = as.Date(datetime),
         tp = as.numeric(tp))

#change units 
nutes<- nutes %>%
  filter(site!="Canal") %>% 
  mutate(doc_uM= (doc*1000000)/(1000*12.011),
         tp_uM= (tp*1000000)/(1000*30.973762),
         nh4_uM= (nh4_n*1000000)/(1000*14.0067),
         no3_uM= (no3_n*1000000)/(1000*14.0067))

tp_plot<- ggplot(nutes,aes(x=date,y=tp_uM, color=site)) +
  geom_point()+
  geom_line(alpha= 0.5) +
  ggtitle("TP")+
  ylab("")+
  facet_wrap(~site , ncol= 3) +
  theme_few()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(legend.position = "none")+
  scale_color_manual(values=c("#35978F", "#8073AC", "#BF812D", "#8C510A","#01665E", "#542788"))+
  scale_shape_manual(values= c(19,1))

tp_plot

doc_plot<- ggplot(nutes,aes(x=date,y=doc_uM, color=site)) +
  geom_point()+
  geom_line(alpha= 0.5) +
  ggtitle("DOC")+
  ylab("")+
  facet_wrap(~site , ncol= 3) +
  theme_few()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(legend.position = "none")+
  scale_color_manual(values=c("#35978F", "#8073AC", "#BF812D", "#8C510A","#01665E", "#542788"))

doc_plot

nh4_plot<- ggplot(nutes,aes(x=date,y=nh4_uM, color=site)) +
  geom_point()+
  geom_line(alpha= 0.5) +
  ggtitle("NH4")+
  ylab("")+
  facet_wrap(~site , ncol= 3) +
  theme_few()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(legend.position = "none")+
  scale_color_manual(values=c("#35978F", "#8073AC", "#BF812D", "#8C510A","#01665E", "#542788"))

nh4_plot

no3_plot<- ggplot(nutes,aes(x=date,y=no3_uM, color=site)) +
  geom_point()+
  geom_line(alpha= 0.5) +
  ggtitle("NO3")+
  ylab("")+
  facet_wrap(~site , ncol= 3) +
  theme_few()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(legend.position = "none")+
  scale_color_manual(values=c("#35978F", "#8073AC", "#BF812D", "#8C510A","#01665E", "#542788"))

no3_plot
```

## Figure 4

First we  need to calculate mean nutrient concentrations because duplicates will create problems when we merge our nutrient and discharge data. 

Note: We are not yet interpolating nutrient concentrations. Right now we just want to know the relationship between observations. 


```{r warning=FALSE, error=FALSE, message=FALSE}
#First get daily q and then combine with Cacao and Yure data
q_daily <- q_preds %>%
  inner_join(raw_cm %>% 
               select(site,datetime,temp)) %>%
  group_by(site,date) %>%
  summarize(q = median(med_q),
            count = n(),
            temp = median(temp)) %>%
  dplyr::filter(count > 4*20)

q_daily<- q_daily %>% 
  select(date, q, site)

cacao_q<- cacao_q %>% 
  mutate(site= "cacao") #add site so can combine data sets

yure_q<- yure_q %>% 
  mutate(site= "yure") #add site so can combine data sets

q_daily<- rbind(cacao_q, yure_q, q_daily) %>% 
  drop_na()

#Now get means of the concentrations (do we have a 1 to 1 match on Qs and nutrient concentrations)

nh4<- nutes %>% 
  select(date, site, nh4_uM) %>% 
  group_by(date, site) %>% 
  mutate(nh4_uM= mean(nh4_uM)) %>% 
  unique()

no3<- nutes %>% 
  select(date, site, no3_uM) %>% 
  group_by(date, site) %>% 
  mutate(no3_uM= mean(no3_uM)) %>% 
  unique()  
  
tp<- nutes %>% 
  select(date, site, tp_uM) %>% 
  group_by(date, site) %>% 
  mutate(tp_uM= mean(tp_uM)) %>% 
  unique() 

doc<- nutes %>% 
  select(date, site, doc_uM) %>% 
  group_by(date, site) %>% 
  mutate(doc_uM= mean(doc_uM)) %>% 
  unique() 
         
nutes1<-cbind(nh4, no3[,3], tp[,3], doc[,3]) #Single obs data set 

#Now combine nutrients and discharge 

complete<- full_join(q_daily, nutes1, by= c("site", "date")) %>% 
  drop_na

complete_long <- complete %>%
  pivot_longer(cols = nh4_uM:doc_uM) %>% 
  mutate(month= month(date)) 

q_v_nutes<-ggplot(complete_long,aes(x=q,y=value, color= site, alpha= month)) + 
  geom_point() + 
  ggthemes::theme_few() + 
  xlab("Q")+
  ylab ("")+
  facet_grid(name~site, scales = 'free')  +
  theme_few()+
  theme(legend.position = "right")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
scale_color_manual(values=c("#35978F", "#8073AC", "#BF812D", "#8C510A","#01665E", "#542788"))

q_v_nutes
```

